### 2024-11-08 teaching/math240f24/draft.tex
lecture 8 nov

\vskip \medskipamount \hbox to\hsize
{\relax \unhcopy \strutbox \cont
{{\tenssbx I. Foundations \kern 3pt}}{2}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 1.\kern .5em Set theory }{3}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 2.\kern .5em Propositional logic}{10}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 3.\kern .5em Predicate logic}{17}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 4.\kern .5em Proofs}{21}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 5.\kern .5em Functions}{28}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 6.\kern .5em Cardinality}{33}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 7.\kern .5em Relations}{37}}
\medskip
\noindent\hbox to\hsize{\relax \unhcopy \strutbox \cont
{{\tenssbx II. Number theory \kern 3pt}}{41}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 8.\kern .5em Division }{42}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 9.\kern .5em Primes }{47}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 10.\kern .5em Modular arithmetic }{51}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 11.\kern .5em Applications of
number theory}{60}}
\medskip
\noindent\hbox to\hsize{\relax \unhcopy \strutbox \cont
{{\tenssbx III. Graph theory \kern 3pt}}{68}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 12.\kern .5em Definitions and
basic notions}{69}}

@@

The calculation above shows that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$
always has $n$ vertices and $\lfloor n^2/4\rfloor$ edges. It is the largest possible number
of edges such that Mantel's theorem does not apply. In other words, from Mantel's theorem alone,
we do not know whether
there must be a triangle in $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$, but adding
a single edge to $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ results in a graph that must
have a triangle by Mantel's theorem.
\bye

Soon we will actually be able to prove that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ does not
contain any triangles. This means that the lower bound of $\lfloor n^2/4\rfloor$ in Mantel's theorem
is the best possible one.

@@

\proclaim Lemma \advthm. A graph $G=(V,E)$ is bipartite if and only if all of its connected components
are bipartite.

>>>>

\vskip \medskipamount \hbox to\hsize
{\relax \unhcopy \strutbox \cont
{{\tenssbx I. Foundations \kern 3pt}}{2}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 1.\kern .5em Set theory }{3}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 2.\kern .5em Propositional logic}{10}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 3.\kern .5em Predicate logic}{17}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 4.\kern .5em Proofs}{21}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 5.\kern .5em Functions}{28}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 6.\kern .5em Cardinality}{33}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 7.\kern .5em Relations}{37}}
\medskip
\noindent\hbox to\hsize{\relax \unhcopy \strutbox \cont
{{\tenssbx II. Number theory \kern 3pt}}{41}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 8.\kern .5em Division }{42}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 9.\kern .5em Primes }{47}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 10.\kern .5em Modular arithmetic }{51}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 11.\kern .5em Applications of
number theory}{60}}
\medskip
\noindent\hbox to\hsize{\relax \unhcopy \strutbox \cont
{{\tenssbx III. Graph theory \kern 3pt}}{68}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 12.\kern .5em Definitions and
basic notions}{69}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 13.\kern .5em Triangles and bipartite
graphs}{74}}
\smallskip
\hbox to\hsize {\relax \unhcopy \strutbox \cont {\hskip10pt 14.\kern .5em Trees}{78}}

@@

The calculation above shows that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$
always has $n$ vertices and $\lfloor n^2/4\rfloor$ edges. It is the largest possible number
of edges such that Mantel's theorem does not apply. In other words, from Mantel's theorem alone,
we do not know whether
there must be a triangle in $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$, but adding
a single edge to $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ results in a graph that must
have a triangle by Mantel's theorem.

We will actually be able to prove that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ does not
contain any triangles. This means that the lower bound of $\lfloor n^2/4\rfloor$ in Mantel's theorem
is the best possible one.

@@

\proclaim\datestamp{08}{xi} Lemma \advthm.
A graph $G=(V,E)$ is bipartite if and only if all of its connected components
are bipartite.
